---
title: "forecasting-HN"
author: "HN"
date: "3/8/2021"
output: html_document
---
```{r}
library(scoringRules)
library(tidyverse)
library(ISOweek) 
library(forecast)
library(fable)
library(tsibble)
library(tsibbledata)
library(lubridate)
library(dplyr)
library(neonstore)
```
Download beetle data
```{r message=FALSE}
## Get the latest beetle target data.  
download.file("https://data.ecoforecast.org/targets/beetles/beetles-targets.csv.gz",
              "beetles-targets.csv.gz")
targets <-  read_csv("beetles-targets.csv.gz")
```
Download Humidity data
```{r}
##Get humidity table
humidity <- neon_read(table = "wss_daily_humid-basic", product = "DP4.00001.001")
```
Summarize humidity data
  note: some weeks don't have humidity data, forcing the min/max fns to produce inf and -inf values, this produces a lot of warning messages when run
```{r}
mean_humidity <- humidity %>%
  mutate(iso_week = ISOweek::date2ISOweek(date)) %>%
  mutate(date = as.Date(date)) %>%
  mutate(time = yearweek(date)) %>%
  group_by(siteID, time) %>%
  summarise(mean_humidity = mean(wssRHMean, na.rm = TRUE), min_humidity = min(wssRHMinimum, na.rm = TRUE), max_humidity = max(wssRHMaximum, na.rm = TRUE))
```
Join humidity and beetle/targets data
```{r}
targets_plusRH <- left_join(targets, mean_humidity, by = c("time", "siteID"))
```

```{r}
targets_ts <- targets_plusRH %>%
  select(siteID, time, richness, abundance, mean_humidity, min_humidity, max_humidity) %>%
  mutate(iso_week = ISOweek::date2ISOweek(time)) %>%
  separate(iso_week, into = c("year", "week", "day"))  %>%
  mutate(time = ISOweek::ISOweek2date(paste(year,week, "1", sep = "-")))

targets_tsb <- as_tsibble(targets_ts, key = siteID, index = time)
```

Check predictors
```{r}
targets_tsb %>%
  GGally::ggpairs(columns = 5:7)
#mean, min, and max are all fairly correlated
#Somewhat arbitrarily selecting meanRH as a predictor variable
```
Training data from years before 2020
```{r}
train <- targets_tsb %>%
  filter(year(time) < 2020) %>%
  select(siteID, time, richness, abundance, mean_humidity, min_humidity, max_humidity)
```

Fit simple linear models
```{r}
fit <- train %>%
  model(mean = TSLM(richness ~ mean_humidity),
        trend = TSLM(richness ~ trend()),
        mix = TSLM(richness ~ mean_humidity + trend())) 
a_fit <- train %>%
  model(mean = TSLM(abundance ~ mean_humidity),
        trend = TSLM(abundance ~ trend()),
        mix = TSLM(abundance ~ mean_humidity + trend()))
```
Forecast using predictors (RH values in 2020)
```{r}
predictors <- targets_tsb %>%
  filter(year(time) > 2019)

fc <- fit %>%
  forecast(new_data = predictors,
           simulate = TRUE, iterations = 50)
a_fc <- a_fit %>%
  forecast(new_data = predictors,
           simulate = TRUE, iterations = 50)
```
CRPS Score
```{r}
 crsp_score <- fc  %>%
  accuracy(targets_tsb, list(crps = CRPS))
a_crsp_score <- a_fc %>%
  accuracy(targets_tsb, list(crps = CRPS))
```
Visualize scores by model
```{r}
crsp_score %>%
  ggplot() +
  geom_boxplot(aes(x = .model, y = crps))
a_crsp_score %>%
  ggplot() +
  geom_boxplot(aes(x = .model, y = crps))
```
Other scores
```{r}
score <- fc  %>%
  accuracy(targets_tsb, measures = list(point_accuracy_measures,interval_accuracy_measures, distribution_accuracy_measures))

score %>%
  group_by(.model) %>%
  summarise(RMSE = mean(RMSE, na.rm = T),
            MAE = mean(MAE, na.rm = T),
            CRPS = mean(CRPS, na.rm = T)) %>%
  arrange(CRPS)
```

Visualize at a given site
```{r}
targets_tsb %>%
  filter(siteID == "CPER") %>%
  autoplot(richness) +
  fc %>% filter(siteID == "CPER") %>% autolayer()
```
Visualize model
```{r}
augment(fit) %>%
  ggplot(aes(x = richness, y = .fitted)) +
  geom_point() +
  labs(
    y = "fitted (predicted values)",
    x = "Data (actual values)",
    title = "Richness"
  ) +
  geom_abline(intercept = 0, slope = 1)
```
