---
title: "introductory forecasting"
author: "Chloe Cho"
date: "2/17/2021"
output: github_document
---

```{r message=FALSE}
library(tidyverse)
library(ISOweek) 
library(forecast)
library(fable)
library(tsibble)
library(tsibbledata)
library(lubridate)
library(dplyr)
```


```{r message=FALSE}
## Get the latest beetle target data.  
download.file("https://data.ecoforecast.org/targets/beetles/beetles-targets.csv.gz",
              "beetles-targets.csv.gz")
targets <-  read_csv("beetles-targets.csv.gz")
```

```{r}
targets_ts <- targets %>%
  select(siteID, time, abundance) %>%
  mutate(iso_week = ISOweek::date2ISOweek(time)) %>%
  mutate(time = as.Date(time)) %>%
  mutate(time = yearmonth(time)) %>% #month year to start...
  group_by(time, siteID) %>%
  summarise(abundance = mean(abundance, na.rm = TRUE))
  #separate(iso_week, into = c("year", "week", "day")) %>%
  # filter(siteID %in% c("DSNY", "OSBS"))

targets_tsb <- as_tsibble(targets_ts, key = siteID, index = time)
targets_tsb
```
### Plot the data (visualize)

```{r}
targets_tsb %>%
  filter(siteID == "ABBY") %>%
  autoplot(abundance) +
  labs(x = "time", title = "Abundance for ABBY")
```

### Define a model (specify)

### Train the model (estimate)
```{r}
train <- targets_tsb %>%
  group_by_key() %>%
  fill_gaps() %>%
  tidyr::fill(abundance, .direction = "down") %>%
  filter(year(time) < 2020)
```


```{r}
fit <- train %>%
  model(NAIVE(abundance), 
        MEAN(abundance),
        SNAIVE(abundance))
fit
```

### Produce forecasts (forecast)
```{r}
fit %>% forecast(h = "1 year")
```

```{r}
fit %>%
  forecast(h = "1 year") %>%
  filter(siteID == "ABBY") %>%
  autoplot(train, level=NULL) +
  labs(x = "time", title = "Abundance for ABBY")
```

### Filling Gaps in Data
"Satellite time series are often affected by permanent gaps like missing observations during winter periods. Often time series methods can not deal with missing observations and require gap-free data. This function fills winter gaps with a constant fill value or according to the approach described in Beck et al. (2006)."

Can either fill with a constant value or a function used to compute fill values. 
https://greenbrown.r-forge.r-project.org/man/FillPermanentGaps.html








