---
title: "time-series-regression-rh"
author: "HN"
date: "3/2/2021"
output: html_document
---

```{r}
library(scoringRules)
library(tidyverse)
library(ISOweek) 
library(forecast)
library(fable)
library(tsibble)
library(tsibbledata)
library(lubridate)
library(dplyr)
library(neonstore)
```

```{r message=FALSE}
## Get the latest beetle target data.  
download.file("https://data.ecoforecast.org/targets/beetles/beetles-targets.csv.gz",
              "beetles-targets.csv.gz")
targets <-  read_csv("beetles-targets.csv.gz")
```
In order to fit the regression, first need to gather the RH data, which occurs from lines 33-96


Get targets file with added summarized RH data
```{r}
# download all of the weather data
neonstore::neon_download(product="DP4.00001.001")
##Get humidity table
humidity <- neon_read(table = "wss_daily_humid-basic", product = "DP4.00001.001")
all_humidities <- humidity %>%
  select(date, siteID, wssRHMean, wssRHMinimum, wssRHMaximum) %>%
  full_join(targets, by = c("date" = "time", "siteID" = "siteID"))
rh_summary <- function(x){
##Select site x'
site <- all_humidities %>%
  filter(siteID  == x)
##Filter for dates at which there is at least one richness or abundance observation
dates <- site %>%
  mutate("abundance.na" = if_else(is.na(abundance), 1, 0), "richness.na" = if_else(is.na(richness), 1, 0)) %>%
  mutate("sum.na" = abundance.na + richness.na) %>%
  filter(sum.na %in% c(0,1)) %>%
  distinct(date)

#Set up vectors to store 2 week info in
meanRH <- NULL
minRH <- NULL
maxRH <- NULL

#For loop that calculates the 2 week info, based on the last 2 weeks from each beetle measurement
for(i in 1:length(dates$date)) {
  a <- filter(site, date %in% c(dates$date[i]-0:13 ))
  meanRH[i] <- mean(a$wssRHMean, na.rm = TRUE)
  minRH[i] <- min(a$wssRHMinimum, na.rm = TRUE)
  maxRH[i] <- max(a$wssRHMaximum, na.rm = TRUE)
}

blan_final <- site %>%
  mutate("abundance.na" = if_else(is.na(abundance), 1, 0), "richness.na" = if_else(is.na(richness), 1, 0)) %>%
  mutate("sum.na" = abundance.na + richness.na) %>%
  filter(sum.na %in% c(0,1)) %>%
  select(siteID, date, richness, abundance) %>%
  bind_cols("meanRH" = meanRH,"minRH" =  minRH,"maxRH"= maxRH)

print(blan_final)
}

compiler <- function() {
  list <- list() # generate an empty list to store each site's df into
sites <- unique(targets$siteID) #get vector of each site
for(i in 1:47) {
  list[[i]] <- rh_summary(sites[i]) # for loop that generates df with RH data for each site
}
df <- NULL # generate empty df to group everything together into
df <- list[[1]]
for(i in 1:46){
  df <- rbind(df, list[[i+1]]) # for loop that adds each site to df
}
targets_rh <- df %>% #replace inf and -inf values with Nan
  mutate("min_inf" = if_else(is.infinite(df$minRH), 0, 1)) %>%
  mutate("max_inf" = if_else(is.infinite(df$maxRH), 0, 1)) %>%
  mutate("min_rh" = minRH * min_inf) %>%
  mutate("max_rh" = maxRH * max_inf) %>%
  select(-c("min_inf", "max_inf", "minRH", "maxRH"))

print(targets_rh)
}
```

Get the 'targets + rh data' df
```{r}
targets_rh <- compiler()
```
Get into correct format
```{r}
targets_ts <- targets_rh %>%
  select(siteID, date, richness, meanRH) %>%
  mutate(iso_week = ISOweek::date2ISOweek(date)) %>%
  separate(iso_week, into = c("year", "week", "day"))  %>%
  mutate(time = ISOweek::ISOweek2date(paste(year,week, "1", sep = "-")))

targets_tsb <- as_tsibble(targets_ts, key = siteID, index = time)
```
Not filling gaps 
```{r}
#targets_tsb <- targets_tsb %>%
#  group_by_key() %>%
#  fill_gaps() %>%
#  tidyr::fill(richness, .direction = "down") 
```
Training data from years before 2020
```{r}
train <- targets_tsb %>%
  filter(year(time) < 2020) %>%
  select(siteID, time, richness, meanRH)
```
Fit simple linear models
```{r}
fit <- train %>%
  model(mean = TSLM(richness ~ meanRH),
        trend = TSLM(richness ~ trend()),
        mix = TSLM(richness ~ meanRH + trend())) 
```
Forecast using predictors (RH values in 2020)
```{r}
predictors <- targets_tsb %>%
  filter(year(time) > 2019)

fc <- fit %>%
  forecast(new_data = predictors,
           simulate = TRUE, iterations = 50)
```
CRPS Score
```{r}
 crsp_score <- fc  %>%
  accuracy(targets_tsb, list(crps = CRPS))
```
Visualize scores by model
```{r}
crsp_score %>%
  ggplot() +
  geom_boxplot(aes(x = .model, y = crps))
```
Visualize Forecast for a given site
```{r}
fc %>% filter(siteID == "CPER") %>%
  autoplot(targets_tsb)
```

